### 2023-05-17 ms.Rmd
Add manuscript skeleton

---
title: ''
shorttitle: 
leftheader: 
author: 
  - name: Matti Vuorre
    affiliation: 1
    corresponding: yes
    address: Department of Social Psychology, Tilburg University
    email: m.j.vuorre@tilburguniversity.edu
affiliation:
  - id: 1
    institution: Department of Social Psychology, Tilburg University
abstract: |

keywords: 
authornote: |
  \noindent \textbf{Working paper. Proceed with caution.}
wordcount: "`r wordcountaddin:::word_count(filename = 'ms.Rmd')`"
bibliography: references.bib
floatsintext: yes
numbersections: false
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
documentclass: apa7
classoption: jou
header-includes:
  - |
    \usepackage{float}
    \setlength{\parskip}{4pt}
output: 
  papaja::apa6_pdf:
    number_sections: true
    keep_tex: false
    highlight: kate
  papaja::apa6_docx: 
    number_sections: true
    keep_tex: false
---

```{r setup, include = FALSE}
# Packages
library(tidyverse)

# Output options
opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  warning = FALSE,
  error = TRUE,
  message = FALSE
)

# Plotting options
theme_set(
  theme_linedraw(base_size = 12) +
    theme(
      strip.background = element_rect(fill = "gray90", color = "gray90"),
      strip.text = element_text(color = "black", hjust = 0),
      panel.grid = element_blank()
    )
)

```

# References {-}

::: {#refs custom-style="Bibliography"}
:::