### 2023-07-14 supplement.qmd
Content

---
subtitle: Supplementary analysis file
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: Contents
    theme: materia
    mainfont: LibertinusSansRegular
    monofont: FiraCodeRegular
    code-fold: true
    # code-link: true # Slow
    code-tools: true
    header-includes: |
      <link rel="stylesheet" media="screen" href="https://fontlibrary.org//face/fira-code" type="text/css"/>
      <link rel="stylesheet" media="screen" href="https://fontlibrary.org//face/libertinus-sans" type="text/css"/> 
---

```{r document-options}
#| include: false
knitr::opts_chunk$set(
  message = FALSE,
  cache = TRUE,
  warning = FALSE,
  echo = TRUE
)
```

This document shows the output of all our supplementary analyses.

# Session info

```{r}
cat(
  "This output from",
  "\nFile:", knitr::current_input(),
  "\nCommit: ", gert::git_commit_id(),
  "\nRepo:", gert::git_remote_info()$url
)
sessionInfo()
```