### 2023-09-01 ms.qmd
Minor code cleaning

---
date: 2023-07-18
bibliography: bibliography.bib
format:
  html: default
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  include = FALSE,
  echo = FALSE
)
```

```{r}
#| label: packages
#| cache: false

library(scales)
library(cmdstanr)
library(brms)
library(ggstance)
library(patchwork)
library(tidyverse)
```

```{r}
#| label: options
#| cache: false

dir.create("models", FALSE)
options(
  mc.cores = 8,
  brms.backend = "cmdstanr"
)
theme_set(
  theme_linedraw(base_size = 10) +
    theme(
      strip.background = element_rect(fill = NA, color = NA),
      strip.text = element_text(colour = "black", hjust = 0),
      axis.text = element_text(size = rel(0.75)),
      panel.grid = element_blank()
    )
)
```

```{r}
#| label: functions

# Format big numbers with a comma
number2 <- function(x, ...) {
  number(x, ..., big.mark = ",")
}

# Replace extreme percentages
percent2 <- function(x, accuracy = .1) {
  x <- percent(x, accuracy = accuracy)
  x <- if_else(x == "100.0%", ">99.9%", x)
  x <- if_else(x == "0.0%", "<0.1%", x)
  x
}
```

```{r}
#| label: data-download

# Download, unzip data from OSF PWS database
path <- "data-raw/data.zip"
if (!file.exists(path)) {
  download.file(
    url = "https://osf.io/download/j48qf/",
    destfile = "data-raw/data.zip"
  )
  dir.create("data", FALSE)
  unzip(
    zipfile = path,
    files = c("data/demographics.csv", "data/study_prompt_answered.csv")
  )
}
```

# Methods

```{r}
#| label: data-load

dat <- read_csv(
  "data/study_prompt_answered.csv",
  col_select = c(
    pid,
    time = Time_utc,
    duration = CurrentSessionLength,
    prompt = LastStudyPromptType,
    mood = response
  )
) |>
  mutate(
    pid = factor(pid),
    mood = mood / 1000
  )

# Calculate raw data summaries
sraw <- list()
sraw$n_res <- dat |>
  count(prompt)
sraw$n_res_total <- number2(sum(sraw$n_res$n))
sraw$n_res <- sraw$n_res |>
  mutate(n = number2(n)) |>
  pivot_wider(names_from = prompt, values_from = n)
sraw$n_sub <- number2(length(unique(dat$pid)))

# Create sessions using heuristics below
dat <- dat |>
  arrange(pid, time) |>
  mutate(
    i = row_number(),
    # Observation is in a new session if this...
    new_session =
      # is the first total observation,
      i == 1 |
      # session duration is zero,
      (duration == 0) |
      # session is shorter than previous session (ticker has reset),
      (duration < lag(duration, default = 0)) |
      # or if more than 30 minutes since previous observation.
      (time > lag(time) + minutes(30)),
    # Session number is the cumulative sum of new sessions...
    session = cumsum(new_session),
    # ... per person
    .by = pid
  ) |>
  select(-c(i, new_session))

# Filter data:
# Include only wellbeing responses that occurred
# in sessions with both "pre" and "post" measures
dat <- dat |>
  filter(prompt == "Wellbeing") |>
  select(-prompt)

# Count of data with all wellbeing observations
dat_total <- dat |>
  summarise(
    n_total = n(),
    n_players = length(unique(pid))
  )

# Include only sessions with "pre-" and "post" measure(s)
dat <- dat |>
  filter(
    # Session has a wellbeing measure at time = 0
    duration[1] == 0,
    # Session has at least two non-missing wellbeing measures 
    # (e.g. pre + 1 post)
    sum(!is.na(mood)) > 1,
    .by = c(pid, session)
  )

# Count of data to be used in these analyses
dat_filtered <- dat |>
  summarise(
    n_total = n(),
    n_players = length(unique(pid))
  )

# Summary
bind_rows(
  "total" = dat_total,
  "filtered" = dat_filtered,
  .id = "Data"
)

# Summaries of filtered data
s <- list()
s$n_sub <- number2(length(unique(dat$pid)))
s$n_obs <- number2(nrow(dat))
s$n_obs_missing <- number2(nrow(filter(dat, is.na(mood))))
s$n_ses <- number2(nrow(count(dat, pid, session)))
```

## Participants

```{r}
#| label: demographics-load

demo <- read_csv(
  "data/demographics.csv",
  col_select = c(pid, country, gender, age)
) |>
  filter(pid %in% unique(dat$pid))

a <- quantile(demo$age, probs = c(.5, .1, .9), na.rm = TRUE)
a <- str_glue("{a[1]} ({a[2]}, {a[3]})")
g <- count(demo, gender, sort = TRUE) |>
  mutate(p = n / sum(n)) |>
  mutate(x = str_glue("{gender} ({number2(n)}, {percent2(p)})"))
n_countries <- length(unique(demo$country[!is.na(demo$country)]))
countries <- count(demo, country, sort = TRUE) |>
  mutate(p = n / sum(n)) |>
  mutate(x = str_glue("{country} ({number2(n)}, {percent2(p)})"))
```

The PowerWash Simulator case study database contains `r sraw$n_res_total` psychological item responses from `r sraw$n_sub` individuals. In this study, we only studied subjective well-being (mood), so did not consider the other five types of item responses in the database (enjoyment, focus, autonomy, competence, and immersion). Therefore, here we used data from `r s$n_sub` participants who provided `r s$n_obs` mood responses (`r s$n_obs_missing` with missing values) from `r s$n_ses` gameplay sessions.

These participants median age (1st & 9th deciles) was `r a`, and four most frequent responses to gender were `r g$x[1:4]`. They hailed from `r n_countries`, with four most represented countries being `r countries$x[1:4]`.

The study procedures were granted ethical approval by Oxford University’s Central University Research Ethics Committee (SSH_OII_CIA_21_011). All participants provided informed consent to participate and reported being 18 years or older.

## Measures

We measured affective well-being with one item.

## Data analysis

```{r}
#| label: data-wrangle-for-analysis

dat <- dat |>
  mutate(
    post = factor(
      row_number() > 1,
      levels = c(FALSE, TRUE),
      labels = c("0", "1")
    ),
    .by = c(pid, session)
  ) |>
  mutate(ps = interaction(pid, session, sep = "_")) |>
  mutate(
    cl = case_when(
      mood == 0 ~ "left",
      mood == 1 ~ "right",
      TRUE ~ "none"
    )
  )
```

Our primary contrast of interest was between mood levels at the beginning of each session and mood levels during the subsequent play session. We modelled each mood report *i* in 1 to `r nrow(dat)` for participant *j* in 1 to `r length(unique(dat$pid))` and participant-session *k* in 1 to `r length(unique(dat$ps))` with a normal distribution censored at the VAS limits [0, 1].

$$
\begin{align*}
\text{mood}_{ijk} &\sim \text{Normal}(\eta_{ijk}, \kappa_{ij}), \\
\eta_{ijk} &= \alpha_0 + \beta_{0j} + \gamma_{0k} + \\ 
  &\quad \ (\alpha_1 + \beta_{1j} + \gamma_{1k})\text{post}_{ijk}, \\
log(\kappa_{ij}) &= \alpha_2 + \beta_{2j} + \alpha_3\text{post}_{ijk}, \\
\begin{bmatrix} \alpha_0 \\ \alpha_1 \\ \alpha_2 \end{bmatrix} &\sim 
  \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 
    \tau_{\alpha_0} &\rho_{{\alpha_0}{\alpha_1}} &\rho_{{\alpha_0}{\alpha_2}} \\ 
    \rho_{{\alpha_0}{\alpha_1}} &\tau_{\alpha_1} &\rho_{{\alpha_1}{\alpha_2}} \\ 
    \rho_{{\alpha_0}{\alpha_2}} &\rho_{{\alpha_1}{\alpha_2}} &\tau_{\alpha_2} 
  \end{bmatrix}\right), \\
  \begin{bmatrix} \gamma_0 \\ \gamma_1 \end{bmatrix} &\sim
    \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \end{bmatrix}, 
    \begin{bmatrix} 
      \tau_{\gamma_0} &\rho_{{\gamma_0}{\gamma_1}} \\ 
      \rho_{{\gamma_0}{\gamma_1}} &\tau_{\gamma_1}
  \end{bmatrix}\right).
\end{align*}
$$

We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package for models [@burknerBrmsPackageBayesian2017].

```{r}
#| label: model-estimate

set.seed(999)
dat_snip <- dat |>
  filter(pid %in% sample(unique(dat$pid), 300))

model <- bf(mood | cens(cl) ~ 0 + post + (0 + post |p| pid) + (1 | ps)) +
  lf(sigma ~ 0 + post + (1 |p| pid)) +
  gaussian()

fit <- brm(
  model,
  data = dat_snip,
  silent = 0,
  threads = 2,
  iter = 1500,
  control = list(adapt_delta = .9),
  file = "models/brm-pre-post-location-scale"
)
summary(fit, prior = TRUE)
```

# Results

## Descriptives

```{r}
#| label: fig-descriptives
#| include: true
#| fig-height: 2.2
#| fig-width: 7
#| fig-cap: '**A**. Histogram of session durations (note log10 x-axis). **B**. Summary of how many sessions each participant completed. **C**. Histogram of all mood ratings.'

p0 <- dat |>
  ggplot() +
  scale_x_continuous(
    expand = expansion(c(0.01, 0.01)),
  ) +
  scale_y_continuous(
    expand = expansion(c(0.001, 0.1)),
  ) +
  geom_histogram(
    col = "white",
    linewidth = .25
  )

dat_session_durations <- dat |>
  filter(duration == max(duration), .by = c(pid, session))

dat_session_counts <- dat |>
  distinct(pid, session) |>
  count(pid, name = "sessions")

p_durations <- p0 %+%
  dat_session_durations +
  aes(duration) +
  scale_x_log10() +
  labs(x = "Session duration (minutes)", y = "Sessions")

p_sessions <- p0 %+%
  dat_session_counts +
  aes(sessions) +
  labs(x = "Sessions", y = "Participants")

p_mood <- p0 +
  aes(mood) +
  labs(x = "Mood", y = "Responses")

(p_durations | p_sessions | p_mood) +
  plot_annotation(tag_level = "A")
```

```{r}
#| label: fig-data
#| include: true
#| fig-width: 7.2
#| fig-height: 5
#| fig-cap: "**A**. Representative data from three participants (rows) showing scatterplots of mood responses over that session's (columns) duration. Each session's first mood report is highlighted in red, subsequent reports are in blue. **B**. Horizontal histograms of all sessions' mood response for that person. **C**. Scatterplot of person-mean mood reports at the beginning of the session (x-axis) and subsequent reports (y-axis). **D**. Histograms of mood ratings at the beginning of session, during the session, and differences between person-means of first and subsequent mood reports."

axislabels <- c(0, 0.5, 1.0)

# Take a quasirandom sample of people who had sufficient number of sessions
# and data per session.
set.seed(99)
tmp <- dat |>
  add_count(pid, session, name = "obs_per_session") |>
  filter(obs_per_session >= 4) |>
  mutate(session = as.numeric(as.factor(session)), .by = pid) |>
  mutate(session_per_person = length(unique(session)), .by = pid) |>
  filter(session_per_person >= 4) |>
  arrange(pid, session) |>
  filter(pid %in% sample(unique(pid), 3)) |>
  mutate(Session = str_glue("Session {session}"), Person = str_glue("Person {fct_anon(pid)}"))

library(ggdist)
p_mood_duration_sessions <- tmp |> 
  ggplot() +
  aes(duration / 60, mood, col = post) +
  scale_color_brewer(
    "Pre-session measure",
    palette = "Set1",
    aesthetics = c("color", "fill")
  ) +
  scale_x_continuous(
    "Session duration (hours)"
  ) +
  scale_y_continuous(
    "Mood",
    limits = c(0, 1),
    breaks = axislabels,
    labels = axislabels
  ) +
  geom_point(size = 1.5, alpha = 1) +
  facet_grid(
    Person ~ Session,
    scales = "free_x"
  ) +
  theme(
    plot.margin = unit(c(8, 8, 8, 8), "pt"),
    # strip.background.y = element_blank(),
    # strip.text.y = element_blank()
  )

p_mood_duration_all <- tmp |> 
  mutate(Session = "All sessions") |> 
  ggplot() +
  aes(y = mood, col = post, fill = post) +
  scale_color_brewer(
    "Pre-session measure",
    palette = "Set1",
    aesthetics = c("color", "fill")
  ) +
  scale_x_continuous(
    "Scaled count",
    expand = expansion(c(0.01, 0.1))
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    # expand = expansion(c(0, 0.01)),
    breaks = axislabels,
    labels = axislabels
  ) +
  coord_fixed(ratio = 0.66) +
  geom_dotplot(
    aes(x = 0),
    dotsize = 0.9,
    binaxis = "y",
    method = "histodot",
    binwidth = 0.06,
    stackgroups = TRUE
  ) +
  facet_grid(
    rows = vars(Person), 
    cols = vars(Session)
  ) +
  theme(
    axis.text = element_blank(), 
    axis.title.y = element_blank(), 
    axis.ticks.y = element_blank(), 
    strip.background.y = element_blank(),
    strip.text.y = element_blank()
    # plot.margin = unit(c(8, 8, 8, 8), "pt")
  )

p_mood_duration <- (p_mood_duration_sessions | p_mood_duration_all) &
  plot_layout(widths = c(80, 20)) &
  theme(legend.position = "none")

tmp <- dat |> 
  summarise(
    mood = mean(mood, na.rm = TRUE),
    .by = c(pid, post)
  ) |> 
  pivot_wider(names_from = post, values_from = mood) |> 
  mutate(delta = `1` - `0`)

p_mood_biscatter <- tmp |> 
  ggplot(aes(`0`, `1`)) +
  scale_x_continuous(
    "Mood at start of session",
    expand = expansion(0.01),
    breaks = axislabels,
    labels = axislabels
  ) +
  scale_y_continuous(
    "Mood during session",
    expand = expansion(0.01),
    breaks = axislabels,
    labels = axislabels
  ) +
  geom_point(
    alpha = .25, size = 1, shape = 1
  ) +
  geom_abline(linewidth = .33, col = "dodgerblue") +
  theme(aspect.ratio = 1)
p_mood_difference <- tmp |> 
  drop_na(delta) |> 
  ggplot(aes(delta)) +
  scale_y_continuous(
    "Participants",
    expand = expansion(c(0.001, 0.1))
  ) +
  scale_x_continuous(
    "Difference (subsequent - first mood)"
  ) +
  coord_cartesian(xlim = c(-0.4, 0.6)) +
  geom_histogram(binwidth = .025, col = "white") +
  stat_summaryh(
    fun.data = mean_cl_boot_h,
    aes(y = 10),
    col = "white",
    fill = "black",
    shape = 21,
    stroke = 1.5, size = 0.75
  )

p_mood_difference <- tmp |> 
  rename(
    `Pre-session` = `0`,
    `During session` = `1`,
    Difference = delta
  ) |> 
  pivot_longer(-pid) |> 
  mutate(name = fct_inorder(name)) |> 
  ggplot(aes(value)) +
  geom_histogram(bins = 30, col = "white") +
  scale_y_continuous(
    "Participants",
    expand = expansion(c(0.001, 0.1))
  ) +
  scale_x_continuous(
    "Mood",
    expand = expansion(c(0.001))
  ) +
  facet_wrap("name", ncol = 3, scales = "free")

(p_mood_duration / (p_mood_biscatter | p_mood_difference)) +
  plot_layout(heights = c(6, 4)) +
  plot_annotation(tag_level = "A")
```

```{r}
#| label: tbl-avg
#| include: true
#| tbl-cap: Population level average mood at beginning and during session, and difference.

hs <- c(
  "Pre" = "post0 = 0",
  "Post" = "post1 = 0",
  "Difference" = "post1 - post0 = 0"
)

h_avg <- hypothesis(fit, hs)$hypothesis
h_avg |> 
  select(` ` = Hypothesis, Estimate, CI.Lower, CI.Upper) |> 
  mutate(across(where(is.numeric), ~percent2(., .01))) |> 
  knitr::kable(digits = 3)
```

```{r}
#| label: wip-contrasts
#| eval: false

h_sub <- hypothesis(fit, hs, scope = "coef", group = "pid")$hypothesis

tmp <- bind_rows(
  h_avg,
  h_sub
)
tmp |> 
  ggplot(aes(Estimate, Hypothesis)) +
  geom_point(
    data = . %>% filter(is.na(Group)),
    size = 1
  ) +
  geom_point(
    data = . %>% filter(!is.na(Group)),
    size = .25
  )

```

# Bibliography

::: {#refs}
:::

# Appendices / supplementary analyses / stuff