### 2023-09-05 ms.qmd
Reparameterize model

---
title: "Mood increases during video game play: A naturalistic observational case study of PowerWash Simulator"
description: Computational notebook supporting our manuscript
author:
  - name: Matti Vuorre
    orcid: 0000-0001-5052-066X
    email: mjvuorre@uvt.nl
    url: https://vuorre.netlify.app
    affiliation:
      - ref: a
      - ref: b
  - name: Nick Ballou
    orcid: 0000-0003-4126-0696
    affiliation:
      - ref: b
  - name: Andrew K. Przybylski
    orcid: 0000-0001-5547-2185
    affiliation:
      - ref: b
affiliations:
  - id: a
    name: Tilburg University
  - id: b
    name: University of Oxford

date: 2023-09-04
bibliography: bibliography.bib
format:
  docx: default
  html: default
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  include = FALSE,
  echo = FALSE
)
```

```{r}
#| label: packages
#| cache: false

library(scales)
library(cmdstanr)
library(brms)
library(ggstance)
library(patchwork)
library(tidyverse)
```

```{r}
#| label: options
#| cache: false

dir.create("models", FALSE)
options(
  mc.cores = 8,
  brms.backend = "cmdstanr"
)
theme_set(
  theme_linedraw(base_size = 10) +
    theme(
      strip.background = element_rect(fill = NA, color = NA),
      strip.text = element_text(colour = "black", hjust = 0),
      axis.text = element_text(size = rel(0.75)),
      panel.grid = element_blank()
    )
)
```

```{r}
#| label: functions

# Format big numbers with a comma
number2 <- function(x, ...) {
  number(x, ..., big.mark = ",")
}

# Replace extreme percentages
percent2 <- function(x, accuracy = .1) {
  x <- percent(x, accuracy = accuracy)
  x <- if_else(x == "100.0%", ">99.9%", x)
  x <- if_else(x == "0.0%", "<0.1%", x)
  x
}
```

```{r}
#| label: data-download

# Download, unzip data from OSF PWS database
path <- "data-raw/data.zip"
if (!file.exists(path)) {
  download.file(
    url = "https://osf.io/download/j48qf/",
    destfile = "data-raw/data.zip"
  )
  dir.create("data", FALSE)
  unzip(
    zipfile = path,
    files = c("data/demographics.csv", "data/study_prompt_answered.csv")
  )
}
```

```{r}
#| label: data-load

dat <- read_csv(
  "data/study_prompt_answered.csv",
  col_select = c(
    pid,
    time = Time_utc,
    duration = CurrentSessionLength,
    prompt = LastStudyPromptType,
    mood = response
  )
) |>
  mutate(
    pid = factor(pid),
    mood = mood / 1000
  )

# Calculate raw data summaries
sraw <- list()
sraw$n_res <- dat |>
  count(prompt)
sraw$n_res_total <- number2(sum(sraw$n_res$n))
sraw$n_res <- sraw$n_res |>
  mutate(n = number2(n)) |>
  pivot_wider(names_from = prompt, values_from = n)
sraw$n_sub <- number2(length(unique(dat$pid)))

# Create sessions using heuristics below
dat <- dat |>
  arrange(pid, time) |>
  mutate(
    i = row_number(),
    # Observation is in a new session if this...
    new_session =
      # is the first total observation,
      i == 1 |
      # session duration is zero,
      (duration == 0) |
      # session is shorter than previous session (ticker has reset),
      (duration < lag(duration, default = 0)) |
      # or if more than 30 minutes since previous observation.
      (time > lag(time) + minutes(30)),
    # Session number is the cumulative sum of new sessions...
    session = cumsum(new_session),
    # ... per person
    .by = pid
  ) |>
  select(-c(i, new_session))

# Filter data:
# Include only wellbeing responses that occurred
# in sessions with both "pre" and "post" measures
dat <- dat |>
  filter(prompt == "Wellbeing") |>
  select(-prompt)

# Count of data with all wellbeing observations
dat_total <- dat |>
  summarise(
    n_total = n(),
    n_players = length(unique(pid))
  )

# Include only sessions with "pre-" and "post" measure(s)
dat <- dat |>
  filter(
    # Session has a wellbeing measure at time = 0
    duration[1] == 0,
    # Session has at least two non-missing wellbeing measures 
    # (e.g. pre + 1 post)
    sum(!is.na(mood)) > 1,
    .by = c(pid, session)
  )

# Count of data to be used in these analyses
dat_filtered <- dat |>
  summarise(
    n_total = n(),
    n_players = length(unique(pid))
  )

# Summary
bind_rows(
  "total" = dat_total,
  "filtered" = dat_filtered,
  .id = "Data"
)

# Summaries of filtered data
s <- list()
s$n_sub <- number2(length(unique(dat$pid)))
s$n_obs <- number2(nrow(dat))
s$n_obs_missing <- number2(nrow(filter(dat, is.na(mood))))
s$n_ses <- number2(nrow(count(dat, pid, session)))
```

# Methods

## Participants

```{r}
#| label: demographics-load

demo <- read_csv(
  "data/demographics.csv",
  col_select = c(pid, country, gender, age)
) |>
  filter(pid %in% unique(dat$pid))

a <- quantile(demo$age, probs = c(.5, .1, .9), na.rm = TRUE)
a <- str_glue("{a[1]} ({a[2]}, {a[3]})")
g <- count(demo, gender, sort = TRUE) |>
  mutate(p = n / sum(n)) |>
  mutate(x = str_glue("{gender} ({number2(n)}, {percent2(p)})"))
n_countries <- length(unique(demo$country[!is.na(demo$country)]))
countries <- count(demo, country, sort = TRUE) |>
  mutate(p = n / sum(n)) |>
  mutate(x = str_glue("{country} ({number2(n)}, {percent2(p)})"))
```

The PowerWash Simulator case study database contains `r sraw$n_res_total` psychological item responses from `r sraw$n_sub` individuals. In this study, we only studied subjective well-being (mood), so did not consider the other five types of item responses in the database (enjoyment, focus, autonomy, competence, and immersion). Therefore, here we used data from `r s$n_sub` participants who provided `r s$n_obs` mood responses (`r s$n_obs_missing` with missing values) from `r s$n_ses` gameplay sessions.

These participants median age (1st & 9th deciles) was `r a`, and four most frequent responses to gender were `r g$x[1:4]`. They hailed from `r n_countries`, with four most represented countries being `r countries$x[1:4]`.

The study procedures were granted ethical approval by Oxford University’s Central University Research Ethics Committee (SSH_OII_CIA_21_011). All participants provided informed consent to participate and reported being 18 years or older.

## Measures

We measured affective well-being with one item.

## Data analysis

```{r}
#| label: data-wrangle-for-analysis

dat <- dat |>
  mutate(
    post = factor(
      row_number() > 1,
      levels = c(FALSE, TRUE),
      labels = c("0", "1")
    ),
    .by = c(pid, session)
  ) |>
  mutate(ps = interaction(pid, session, sep = "_")) |>
  mutate(
    cl = case_when(
      mood == 0 ~ "left",
      mood == 1 ~ "right",
      TRUE ~ "none"
    )
  )
```

Our primary contrast of interest was between player mood at the beginning of each session and mood during the subsequent play session. We implemented this contrast within a censored three-level hierarchical location-scale model: First, because mood was measured with a VAS, we specified a censored gaussian model of the responses. We then modelled the mean of the response distribution on indicator variables for time (pre, post), and allowed both parameters to vary randomly across players. In addition, we modelled random average moods over player-sessions. Also, because equal variances across people in such naturalistic observation seems unlikely, we specified a similar model for the response distribution's standard deviation, but only modelled random deviations across participants, and so assumed equal residual variation across sessions within players.

More formally, we modelled each mood response *i* in 1 to `r s$n_obs` from participant *j* in 1 to `r s$n_sub` and participant-session *k* in 1 to `r s$n_ses` with a normal distribution censored at the VAS limits [0, 1]. We then modelled the distribution's mean on indicator variables (0 or 1) for pre-play and during play, their random per-player deviations $\beta$, and a random deviation for player-sessions $\gamma_0$. We modelled the residual deviation on a global average and per-player deviations, through a log link function:

$$
\begin{align*}
\text{mood}_{ijk} &\sim \text{CensNormal}(\eta_{ijk}, \kappa_{ij}, 0, 1), \\
\eta_{ijk} &= (\alpha_0 + \beta_{0j})\text{pre}_{ijk} +
  (\alpha_1 + \beta_{1j})\text{post}_{ijk} +
  \gamma_{0k}, \\
log(\kappa_{ij}) &= \alpha_2 + \beta_{2j}, \\
\begin{bmatrix} \alpha_0 \\ \alpha_1 \\ \alpha_2 \end{bmatrix} &\sim 
  \text{MVN}\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 
    \tau_{\alpha_0} &\rho_{{\alpha_0}{\alpha_1}} &\rho_{{\alpha_0}{\alpha_2}} \\ 
    \rho_{{\alpha_0}{\alpha_1}} &\tau_{\alpha_1} &\rho_{{\alpha_1}{\alpha_2}} \\ 
    \rho_{{\alpha_0}{\alpha_2}} &\rho_{{\alpha_1}{\alpha_2}} &\tau_{\alpha_2} 
  \end{bmatrix}\right), \\
  \gamma_0 &\sim \text{Normal}(0, \tau_{\gamma_0}).
\end{align*}
$$

```{r}
#| label: model-estimate

contrasts(dat$post) <- c(-0.5, 0.5)

model <- bf(mood ~ 1 + post + (1 + post | pid) + (1 | ps)) +
  gaussian()

fit <- brm(
  model,
  data = dat,
  silent = 0, # Required for printing mcmc progress on some systems
  threads = 2,
  iter = 500,
  control = list(adapt_delta = .95),
  file = "models/brm-pre-post-pid-ps"
)
```

```{r}
#| label: model-convergence
#| eval: false

hist(rhat(fit))
plot(fit)
pairs(fit$fit, pars = variables(fit)[1:7])
```

We analyzed the data with R [@rcoreteamLanguageEnvironmentStatistical2023] and used the brms package to estimate and post-process the models [@burknerBrmsPackageBayesian2017]. We used Stan's HMC sampler to draw `r number2(ndraws(fit))` samples from the posterior distribution, and used brms' default prior distributions on all parameters, and used numerical and graphical checks to ensure model convergence.

# Results

## Descriptives

```{r}
#| label: fig-descriptives
#| include: true
#| fig-height: 2.2
#| fig-width: 7
#| fig-cap: '**A**. Histogram of session durations (note log10 x-axis). **B**. Summary of how many sessions each participant completed. **C**. Histogram of all mood ratings.'

p0 <- dat |>
  ggplot() +
  scale_x_continuous(
    expand = expansion(c(0.01, 0.01)),
  ) +
  scale_y_continuous(
    expand = expansion(c(0.001, 0.1)),
  ) +
  geom_histogram(
    col = "white",
    linewidth = .25
  )

dat_session_durations <- dat |>
  filter(duration == max(duration), .by = c(pid, session))

dat_session_counts <- dat |>
  distinct(pid, session) |>
  count(pid, name = "sessions")

p_durations <- p0 %+%
  dat_session_durations +
  aes(duration) +
  scale_x_log10() +
  labs(x = "Session duration (minutes)", y = "Sessions")

p_sessions <- p0 %+%
  dat_session_counts +
  aes(sessions) +
  labs(x = "Sessions", y = "Participants")

p_mood <- p0 +
  aes(mood) +
  labs(x = "Mood", y = "Responses")

(p_durations | p_sessions | p_mood) +
  plot_annotation(tag_level = "A")
```

```{r}
#| label: fig-data
#| include: true
#| fig-width: 7.2
#| fig-height: 5
#| fig-cap: "**A**. Representative data from three participants (rows) showing scatterplots of mood responses over that session's (columns) duration. Each session's first mood report is highlighted in red, subsequent reports are in blue. **B**. Horizontal histograms of all sessions' mood response for that person. **C**. Scatterplot of person-mean mood reports at the beginning of the session (x-axis) and subsequent reports (y-axis). **D**. Histograms of mood ratings at the beginning of session, during the session, and differences between person-means of first and subsequent mood reports."

axislabels <- c(0, 0.5, 1.0)

# Take a quasirandom sample of people who had sufficient number of sessions
# and data per session.
set.seed(99)
tmp <- dat |>
  add_count(pid, session, name = "obs_per_session") |>
  filter(obs_per_session >= 4) |>
  mutate(session = as.numeric(as.factor(session)), .by = pid) |>
  mutate(session_per_person = length(unique(session)), .by = pid) |>
  filter(session_per_person >= 4) |>
  arrange(pid, session) |>
  filter(pid %in% sample(unique(pid), 3)) |>
  mutate(Session = str_glue("Session {session}"), Person = str_glue("Person {fct_anon(pid)}"))

library(ggdist)
p_mood_duration_sessions <- tmp |> 
  ggplot() +
  aes(duration / 60, mood, col = post) +
  scale_color_brewer(
    "Pre-session measure",
    palette = "Set1",
    aesthetics = c("color", "fill")
  ) +
  scale_x_continuous(
    "Session duration (hours)"
  ) +
  scale_y_continuous(
    "Mood",
    limits = c(0, 1),
    breaks = axislabels,
    labels = axislabels
  ) +
  geom_point(size = 1.5, alpha = 1) +
  facet_grid(
    Person ~ Session,
    scales = "free_x"
  ) +
  theme(
    plot.margin = unit(c(8, 8, 8, 8), "pt"),
    # strip.background.y = element_blank(),
    # strip.text.y = element_blank()
  )

p_mood_duration_all <- tmp |> 
  mutate(Session = "All sessions") |> 
  ggplot() +
  aes(y = mood, col = post, fill = post) +
  scale_color_brewer(
    "Pre-session measure",
    palette = "Set1",
    aesthetics = c("color", "fill")
  ) +
  scale_x_continuous(
    "Scaled count",
    expand = expansion(c(0.01, 0.1))
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    # expand = expansion(c(0, 0.01)),
    breaks = axislabels,
    labels = axislabels
  ) +
  coord_fixed(ratio = 0.66) +
  geom_dotplot(
    aes(x = 0),
    dotsize = 0.9,
    binaxis = "y",
    method = "histodot",
    binwidth = 0.06,
    stackgroups = TRUE
  ) +
  facet_grid(
    rows = vars(Person), 
    cols = vars(Session)
  ) +
  theme(
    axis.text = element_blank(), 
    axis.title.y = element_blank(), 
    axis.ticks.y = element_blank(), 
    strip.background.y = element_blank(),
    strip.text.y = element_blank()
    # plot.margin = unit(c(8, 8, 8, 8), "pt")
  )

p_mood_duration <- (p_mood_duration_sessions | p_mood_duration_all) &
  plot_layout(widths = c(80, 20)) &
  theme(legend.position = "none")

tmp <- dat |> 
  summarise(
    mood = mean(mood, na.rm = TRUE),
    .by = c(pid, post)
  ) |> 
  pivot_wider(names_from = post, values_from = mood) |> 
  mutate(delta = `1` - `0`)

p_mood_biscatter <- tmp |> 
  ggplot(aes(`0`, `1`)) +
  scale_x_continuous(
    "Mood at start of session",
    expand = expansion(0.01),
    breaks = axislabels,
    labels = axislabels
  ) +
  scale_y_continuous(
    "Mood during session",
    expand = expansion(0.01),
    breaks = axislabels,
    labels = axislabels
  ) +
  geom_point(
    alpha = .25, size = 1, shape = 1
  ) +
  geom_abline(linewidth = .33, col = "dodgerblue") +
  theme(aspect.ratio = 1)
p_mood_difference <- tmp |> 
  drop_na(delta) |> 
  ggplot(aes(delta)) +
  scale_y_continuous(
    "Participants",
    expand = expansion(c(0.001, 0.1))
  ) +
  scale_x_continuous(
    "Difference (subsequent - first mood)"
  ) +
  coord_cartesian(xlim = c(-0.4, 0.6)) +
  geom_histogram(binwidth = .025, col = "white") +
  stat_summaryh(
    fun.data = mean_cl_boot_h,
    aes(y = 10),
    col = "white",
    fill = "black",
    shape = 21,
    stroke = 1.5, size = 0.75
  )

p_mood_difference <- tmp |> 
  rename(
    `Pre-session` = `0`,
    `During session` = `1`,
    Difference = delta
  ) |> 
  pivot_longer(-pid) |> 
  mutate(name = fct_inorder(name)) |> 
  ggplot(aes(value)) +
  geom_histogram(bins = 30, col = "white") +
  scale_y_continuous(
    "Participants",
    expand = expansion(c(0.001, 0.1))
  ) +
  scale_x_continuous(
    "Mood",
    expand = expansion(c(0.001))
  ) +
  facet_wrap("name", ncol = 3, scales = "free")

(p_mood_duration / (p_mood_biscatter | p_mood_difference)) +
  plot_layout(heights = c(6, 4)) +
  plot_annotation(tag_level = "A")
```

```{r}
#| label: tbl-avg
#| include: true
#| tbl-cap: Population level average mood at beginning and during session, and difference.

hs <- c(
  "Pre" = "post0 = 0",
  "Post" = "post1 = 0",
  "Difference" = "post1 - post0 = 0"
)

h_avg <- hypothesis(fit, hs)$hypothesis
h_avg |> 
  select(` ` = Hypothesis, Estimate, CI.Lower, CI.Upper) |> 
  mutate(across(where(is.numeric), ~percent2(., .01))) |> 
  knitr::kable(digits = 3)
```

```{r}
#| label: wip-contrasts
#| eval: false

h_sub <- hypothesis(fit, hs, scope = "coef", group = "pid")$hypothesis

tmp <- bind_rows(
  h_avg,
  h_sub
)
tmp |> 
  ggplot(aes(Estimate, Hypothesis)) +
  geom_point(
    data = . %>% filter(is.na(Group)),
    size = 1
  ) +
  geom_point(
    data = . %>% filter(!is.na(Group)),
    size = .25
  )

```

# Bibliography

::: {#refs}
:::

# Appendices / supplementary analyses / stuff