### 2023-09-07 continuous-time-draft.qmd
draft

---
title: "Continuous time"
date: now
author: Kristoffer Magnusson
format:
  html:
    code-fold: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  include = TRUE,
  echo = TRUE
)
```

```{r}
#| label: packages
#| cache: false
library(patchwork)
library(tidyverse)
library(lme4)
library(knitr)
library(emmeans)
```

Data cleaning copied from `ms.qmd` to create sessions and the pre/post contrast.
```{r}
#| label: options
#| cache: false
dat <- read_rds("data/session.rds")
dat2 <- dat |>
  filter(prompt == "Wellbeing") |>
  select(-prompt)

dat2 <- dat2 |>
  filter(
    # Session has a wellbeing measure at time = 0
    duration[1] == 0,
    # Session has at least two non-missing wellbeing measures 
    # (e.g. pre + 1 post)
    sum(!is.na(mood)) > 1,
    .by = c(pid, session)
  )

dat2 <- dat2 |>
  mutate(
    post = factor(
      row_number() > 1,
      levels = c(FALSE, TRUE),
      labels = c("0", "1")
    ),
    .by = c(pid, session)
  ) |>
  mutate(ps = interaction(pid, session, sep = "_")) |>
  mutate(
    cl = case_when(
      mood == 0 ~ "left",
      mood == 1 ~ "right",
      TRUE ~ "none"
    )
  )
```

## Exploring sessions
Let's look at one participant and compare the pre/post data set to the complete data.

```{r}
#| label: fig-duration
#| fig-cap: Comparing full data (top) with pre-post (bottom) data
p0 <- ggplot(
  filter(dat, 
    pid == "p6622",
    session %in% 14:16
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(prompt == "Wellbeing")),
    size = 2
    ) +
  scale_x_datetime(limits = c(ymd_hms("2022-09-19 16:00:00"), ymd_hms("2022-09-20 00:00:00"))) +
  scale_y_continuous(breaks = 14:16) +
  facet_wrap(~pid)

p1 <- ggplot(
  filter(dat2, 
    pid == "p6622",
    session %in% 14:16
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(post)),
    size = 2
    ) +
  scale_x_datetime(limits = c(ymd_hms("2022-09-19 16:00:00"), ymd_hms("2022-09-20 00:00:00"))) +
  scale_y_continuous(breaks = 14:16) +
  facet_wrap(~pid)
p0 / p1
```

## Concerns with the pre/post contrast.
There are issues with the pre/post contrasts.

### 1. Session duration doesn't always start at 0
As can be seen in @fig-duration, session 16 isn't included in the pre/post data, even tho the first response is a WB prompt. This is because it's duration is 6. 
```{r}
filter(dat, 
  pid == "p6622",
  session %in% 14:16
) %>% 
kable()

filter(dat2, 
  pid == "p6622",
  session %in% 14:16
) %>% 
kable()
```

### 2. Lot's of data is discarded
As can be seen in the @fig-pre-post there can be lot's of data that is not covered by the pre-post contrast. This might not be a problem as the prompts were randomized. But it's a lot of PWS gameplay that's also ignored.

```{r}
#| label: fig-pre-post
#| fig-cap: Comparing full data (top) with pre-post (bottom) data
p0 <- ggplot(
  filter(dat, 
    pid == "p10",
    session %in% 16:26
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(prompt == "Wellbeing")),
    size = 2
    ) +
  scale_x_datetime(
    limits = c(
      ymd_hms("2023-02-25 00:00:00"), 
      ymd_hms("2023-02-29 00:00:00"
      )
    )
  ) +
  scale_y_continuous(breaks = 16:26)

p1 <- ggplot(
  filter(dat2, 
    pid == "p10",
    session %in% 16:26
    ),
  aes(
    time,
    session,
    group = session
  )
) + 
  geom_point(
    aes(color = factor(post)),
    size = 2
    ) +
  scale_x_datetime(
    limits = c(
      ymd_hms("2023-02-25 00:00:00"), 
      ymd_hms("2023-02-29 00:00:00")
      )
    ) +
  scale_y_continuous(breaks = 16:26)
p0 / p1
```

### 3. There are multiple post responses within a session
The "post" contrast is the average mood during a session. I know this is intentional, but I'm not sure if it's ideal. I think continuous time would be more interesting. The contrast averages over sessions that are too irregular, and will be hard to interpret. A model with continuous time within sessions would solve this.

```{r}
filter(dat, pid == "p10", session == 16) %>% 
  kable()

filter(dat2, pid == "p10", session == 16) %>% 
  kable()
```

### 4. The post contrast doesn't represent the whole session
This can also be seen in the data above. The last post measure in session 16 is at 2023-02-25 11:31:43, but the actual session continues for at least an hour longer, as can be seen in the full data: 2023-02-25 12:41:00. I'm not sure if this matter, but it feels weird that this isn't captured by the model. I guess this is related to concern #2, I think we need to somehow encode the "dose" of the PWS "treatment", and the planned missingness of the mood prompts.

### The post contrast could be replaced with continuous time
A model of the form

```r
mood ~ spline(time, knots) + (1 | session) + (1 | id)
```

would allow modeling change during session while still accounting for the highly irregular sessions.

There is also the option to scale time to 1 on a session level to more closely mimic a pre/post design. This would have the advantage of being able to include all wellbeing responses, regardless if a session starts with a wellbeing measure.

### Fit model
Reset duration within each session
```{r}
dat <- dat %>% 
  mutate(
    duration = duration - duration[1],
    hours = duration / 60,
    .by = c(pid, session)
  ) %>% 
  filter(prompt == "Wellbeing")
```

Fit linear change
```{r}
library(rms)
fit0 <- lmer(mood ~ hours + (1 | session) + (1 | pid), data = dat)
summary(fit0)
```

What's the distribution of session lengths
```{r}
median_ses_len <- dat %>% 
  group_by(pid, session) %>% 
  summarise(session_length = max(duration)) %>% 
  group_by(pid) %>% 
  summarise(median_len = median(session_length))

median(median_ses_len$median_len)

mean(median_ses_len$median_len > 60)

```

"Typical" session length is 24 minutes. Few participants typically play for more than an hour. Plot estimated trend over 0 to 2 hours.

```{r}
emm0 <- emmeans(
  fit0, 
  ~hours, 
  at = list(hours = seq(0,2, length.out = 100)),
  lmer.df = "asymptotic"
  )

p0 <- as.data.frame(emm0) %>% 
  ggplot(aes(hours, emmean)) + 
  geom_line() +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
  labs(x = "Session duration (hours)")
```

Let's try non-linear change using `rms::rcs` with 5 knots with default placement, which should place most knots during the first hour.
```{r}
attr(rcs(dat$hours, 5), "parms")
```

Fit model with only random intercepts.
```{r}
fit1 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 | pid), data = dat)
emm1 <- emmeans(
  fit1, 
  ~hours, 
  at = list(hours = seq(0,2, length.out = 100)),
  lmer.df = "asymptotic"
  )
p1 <- as.data.frame(emm1) %>% 
  ggplot(aes(hours, emmean)) + 
  geom_line() +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
  labs(x = "Session duration (hours)")
```

Adding `rcs(hours, 5)` as random effects would be messy, so let's fit the model with a linear random slope at the participant level. I don't think random slopes at the session level is needed? Moreover, I would expect that the exact random slopes specification have little impact on fixed effects in this case.
```{r}
path <- "models/lmm-fit2.Rds"
if (file.exists(path)) {
  fit2 <- read_rds(path)
} else {
  fit2 <- lmer(mood ~ rcs(hours, 5) + (1 | session) + (1 + hours | pid), data = dat)
  write_rds(fit2, path)
}
emm2 <- emmeans(
  fit2, 
  ~hours, 
  at = list(hours = seq(0,2, length.out = 100)),
  lmer.df = "asymptotic"
  )
p2 <- as.data.frame(emm2) %>% 
  ggplot(aes(hours, emmean)) + 
  geom_line() +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL), alpha = 0.25) +
  labs(x = "Session duration (hours)")
```

Plots models, ignore model 2 as it's basically identical to model 1.

```{r}
p0 / p1
```

I think the shape of the spline makes sense.

TODO:

- GAMM?
- Is censoring really needed?
- Visualize the % that have positive vs negative change over a session? or over 30 / 60 minutes.