### 2023-09-19 wrangle.qmd

---
title: Wrangle
---

This notebook wrangles the raw data on OSF to something more suitable for our current analyses.

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  include = TRUE,
  echo = TRUE
)
```

```{r}
#| label: packages
#| cache: false

library(tidyverse)
source("functions.R")
```

```{r}
#| label: data-download

# Download, unzip data from OSF PWS database
path <- "data-raw/data.zip"
if (!file.exists(path)) {
  dir.create("data-raw", FALSE)
  download.file(
    url = "https://osf.io/download/j48qf/",
    destfile = "data-raw/data.zip"
  )
  dir.create("data", FALSE)
  unzip(
    zipfile = path,
    files = c("data/demographics.csv", "data/study_prompt_answered.csv")
  )
}
```

```{r}
#| label: data-load

dat <- read_csv(
  "data/study_prompt_answered.csv",
  col_select = c(
    pid,
    time = Time_utc,
    duration = CurrentSessionLength,
    prompt = LastStudyPromptType,
    mood = response
  )
) |>
  mutate(
    pid = factor(pid),
    mood = mood / 1000
  )

# Calculate raw data summaries
sraw <- list()
sraw$n_res <- dat |>
  count(prompt)
sraw$n_res_total <- number2(sum(sraw$n_res$n))
sraw$n_res <- sraw$n_res |>
  mutate(n = number2(n)) |>
  pivot_wider(names_from = prompt, values_from = n)
sraw$n_sub <- number2(length(unique(dat$pid)))

# Create sessions using heuristics below
dat <- dat |>
  arrange(pid, time) |>
  mutate(
    i = row_number(),
    # Observation is in a new session if this...
    new_session =
      # is the first total observation,
      i == 1 |
      # session duration is zero,
      (duration == 0) |
      # session is shorter than previous session (ticker has reset),
      (duration < lag(duration, default = 0)) |
      # or if more than 30 minutes since previous observation.
      (time > lag(time) + minutes(30)),
    # Session number is the cumulative sum of new sessions...
    session = cumsum(new_session),
    # ... per person
    .by = pid
  ) |>
  select(-c(i, new_session))

# Filter data:
# Include only wellbeing responses that occurred
# in sessions with both "pre" and "post" measures
dat <- dat |>
  filter(prompt == "Wellbeing") |>
  select(-prompt)

# Count of data with all wellbeing observations
dat_total <- dat |>
  summarise(
    n_total = n(),
    n_players = length(unique(pid))
  )

# Include only sessions with "pre-" and "post" measure(s)
dat <- dat |>
  filter(
    # Session has a wellbeing measure at time = 0
    duration[1] == 0,
    # Session has at least two non-missing wellbeing measures 
    # (e.g. pre + 1 post)
    sum(!is.na(mood)) > 1,
    .by = c(pid, session)
  )

# Count of data to be used in these analyses
dat_filtered <- dat |>
  summarise(
    n_total = n(),
    n_players = length(unique(pid))
  )

# Summary
bind_rows(
  "total" = dat_total,
  "filtered" = dat_filtered,
  .id = "Data"
)

# Summaries of filtered data
s <- list()
s$n_sub <- number2(length(unique(dat$pid)))
s$n_obs <- number2(nrow(dat))
s$n_obs_missing <- number2(nrow(filter(dat, is.na(mood))))
s$n_ses <- number2(nrow(count(dat, pid, session)))
```

```{r}
dir.create("data", FALSE)
saveRDS(dat, "data/data.rds")
saveRDS(s, "data/s.rds")
saveRDS(sraw, "data/sraw.rds")
```