### 2023-10-06 continuous-time.qmd

# Continuous time {#sec-continuous-time}

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  include = TRUE,
  echo = TRUE
)
```

In the main analysis in @sec-analyze we focused on contrasting the mood at the beginning of a session with the average mood during session. In this document we focus on two additional research questions:

1. How does mood change during a session?
2. Is change during a session related to the participants' pretest mood?

To answer these questions time is modeled as a continuous variable.

## Setup
```{r}
#| label: packages
#| cache: false

library(scales)
library(cmdstanr)
library(posterior)
library(brms)
library(patchwork)
library(tidyverse)
library(splines)
library(emmeans)
library(lme4)
library(broom.mixed)
library(knitr)
library(rstanarm)
source("R/functions.R")
source("R/fit_cached.R")
source("R/common.R")
```

## Data
Load the data.
```{r}
#| label: load-data

dat <- read_rds("data/data.rds")

if (!is.na(N_SUBSET)) {
  pid_subset <- sample(unique(dat$pid), N_SUBSET)
  dat <- filter(dat, pid %in% pid_subset) |> 
    droplevels()
}
```

## Main model
Define a spline that uses 4 degrees of freedom. We base this on the quantiles of `hours`.

```{r}
#| label: spline-percentiles
quantile(dat$hours)
```

Not all participants contributed an equal number of observations. We'll calculate the max session length per participant.

```{r}
#| label: tbl-session-max-prop
#| tbl-cap: Proportion of participants that had at least one session last longer than gt_* hours
dat |>
  summarise(
    max = max(hours),
    .by = c(pid)
  ) |>
  summarise(
    gt_0.18 = mean(max > 0.18),
    gt_0.51 = mean(max > 0.51),
    gt_1.3 = mean(max > 1.3),
    gt_2 = mean(max > 2),
  ) |>
  kable(digits = 2)
```

@tbl-session-max-prop shows that 85% of the participant have at least one session that lasted for over 0.51 hours, and 64% for over 1.3 hours. So we will base our knot placement on these quantiles.

```{r}
#| label: define-spline
#| code-fold: show
ns4 <- function(x) {
  ns(
    x,
    knots = c(0.18, 0.5, 1.3),
    Boundary.knots = c(0, 3)
  )
}
```

The main model is a LMM, we include random effects for each piece of the spline and a random intercept at the session level.

```{r}
#| label: lmm-ns4
#| code-fold: show
fit <- fit_cached(
  "models/lmm-ns4.Rds",
  lmer(
    mood ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
    data = dat
  )
)
```

This model included `r number2(ngrps(fit)["pid"])` participants, with a total of `r number2(ngrps(fit)["ps"])` sessions and `r number2(nobs(fit))` observations.

The spline is plotted in @fig-plot-curve, and we can see that there's a small but sharp increase early during a session.

```{r}
#| label: fig-plot-curve
#| fig-cap: Estimated mood during a play session
#| fig-height: 3
#| fig-width: 8

hours <- seq(0, 3, length.out = 100)
emm1 <- emmeans(
  fit,
  ~hours,
  at = list(hours = hours),
  lmer.df = "asymptotic"
)
p_curve <- as.data.frame(emm1) |>
  ggplot(aes(hours, emmean)) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 1)"
  ) +
  scale_x_hours
p_curve
```

@fig-plot-change shows the same spline, but as a contrast where hour = 0 is the reference value.

```{r}
#| label: fig-plot-change
#| fig-cap: Estimated difference in mood over a session compared to the start of a session
#| fig-height: 3
#| fig-width: 8

d_change <- emmeans::contrast(
  emm1,
  method = "trt.vs.ctrl",
  ref = "hours0"
) |>
  confint() |>
  as.data.frame() |>
  mutate(
    hours = hours[-1]
  )

p_change <- d_change |>
  ggplot(aes(hours, estimate)) +
  scale_y_continuous(
    expand = expansion(c(0.1, 0.1))
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL
    ),
    alpha = 0.25
  ) +
  labs(x = "Session duration (hours)") +
  scale_x_hours
p_change
```

@tbl-change-contrast shows the estimated change from baseline after 15min, 30 min, 1 hour, 2 hours, and 3 hours.
```{r}
#| label: tbl-change-contrast
#| tbl-cap: Estimated change from baseline
emmeans(
  fit,
  ~hours,
  at = list(hours = c(0, 15 / 60, 0.5, 1, 2, 3)),
  lmer.df = "asymptotic"
) |>
  emmeans::contrast(
    method = "trt.vs.ctrl",
    ref = "hours0",
  ) |>
  confint() |>
  kable(
    digits = 3
  )
```

### Relationship between change and mood at the start of a session
We are also interested in the relationship between the initial mood and change over time. As the intercepts are correlated with the random slopes we can plot the implied splines for someone with mood levels 0, 1, and 2 SD below the average at the start of a session. The logic here is simply: if cor(intercept, slope) = 0.5, then someone with an intercept equal to -1SD is expected to have a slope equal to -0.5 * sd(slope). This is shown in @fig-intercept-sds, which tells us that someone with a lower mood at the beginning of a session is expected to have their mood increase more during a session.

Interpret this with caution.

```{r}
#| label: fig-intercept-sds
#| fig-cap: Estimated curves for participants with an intercept equal to 0, -1SD, and -2SD.
#| fig-height: 3
#| fig-width: 8
source("R/get_trend.R")

pre_labels <- c("-2 SD", "-1 SD", "Mean", "1 SD")
pre_sds <- c(-2, -1, 0, 1)
colors <- c(
  "1 SD" = "#4daf4a",
  "-1 SD" = "#e41a1c",
  "-2 SD" = "#377eb8",
  "Mean" = "#000000"
)
color_scale <- list(
  scale_color_manual(
    "Pre",
    values = colors
  ),
    scale_fill_manual(
    "Pre",
    values = colors
  )
)

p_int_sd <- map(
  pre_sds,
  \(sd) get_trend(fit, dat, sd = sd)
) |>
  bind_rows() |>
  mutate(
    intercept = factor(
      rep(
        pre_labels,
        each = 100
      ),
      levels = pre_labels
    )
  ) |>
  ggplot(
    aes(
      x,
      y,
      color = intercept,
      group = intercept
    )
  ) +
  geom_line() +
  labs(
    x = "Session duration (hours)",
    y = "Mood (0 - 100)"
  ) +
  color_scale +
  scale_x_hours

p_int_sd
```

The curves in @fig-intercept-sds are easier to compare if we plot them without an intercept, as can be seen in @fig-no-intercept.

```{r}
#| label: fig-no-intercept
#| fig-cap: Estimated change curves for participants with an intercept equal to 0, -1SD, and -2SD.
#| fig-height: 3
#| fig-width: 8
p_int_sd_change <- map(
  pre_sds,
  \(sd) get_trend(
    fit,
    dat,
    sd = sd,
    intercept = FALSE
  )
) |>
  bind_rows() |>
  mutate(
    intercept = factor(
      rep(
        pre_labels,
        each = 100
      ),
      levels = pre_labels
    )
  ) |>
  ggplot(
    aes(
      x,
      y,
      color = intercept,
      group = intercept
    )
  ) +
  geom_line() +
  geom_ribbon(
    data = d_change,
    aes(
      x = hours,
      y = NULL,
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      group = "0",
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  labs(
    x = "Session duration (hours)",
    y = "Change in Mood (0 - 100)"
  ) +
  color_scale +
  scale_x_hours
p_int_sd_change
```

### Pretest x time continuous interaction
Modeling the relationship between baseline wellbeing and change using random effects is limited. Here we'll fit a separate interaction model to the subset of participants that began their session by answering a mood question. This allows us to explicitly model the interaction between the pretest values and the within-session change, without having to infer this from the multivariate normal distribution of the random effects. This also allows us to model this as a non-linear relationship.

```{r}
#| label: lmer-pre-interaction
#| code-fold: show
dat_pre <- dat |>
  filter(
    # Session has a wellbeing measure at time = 0
    hours[1] == 0,
    .by = c(pid, session)
  ) |>
  mutate(
    pre = mood[1],
    .by = c(pid, session)
  ) |>
  filter(
    hours > 0
  ) |> 
  droplevels()

fit3 <- fit_cached(
  "models/lmm-pre-interaction.Rds",
  lmer(
    mood ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps),
    data = dat_pre
  )
)
```

This model included `r number2(ngrps(fit3)["pid"])` participants, with a total of `r number2(ngrps(fit3)["ps"])` sessions and `r number2(nobs(fit3))` observations.

Next we'll calculate the pre values we are interested in plotting.

```{r}
#| label: tbl-pre-sum
#| tbl-cap: Average mood levels the start of a session
pre_sum <- dat_pre |>
  summarise(
    pre_mean = median(pre),
    .by = pid
  ) |>
  summarise(
    mean = mean(pre_mean),
    sd = sd(pre_mean)
  )
kable(pre_sum, digits = 2)
```

However, baseline mood is not normally distributed, so we'll use percentiles instead. 
```{r}
#| label: tbl-pre-sd-values
#| tbl-cap: Distribution of average mood levels at the start of a session
#pre_values <- pre_sum$mean + pre_sds * pre_sum$sd

pre <- dat_pre |> 
  summarise(pre = pre[1], .by = ps) |> 
  pull(pre)
pre_values <- quantile(pre, c(0.05, .25, 0.5, 0.75))
pre_labels_x <- c("5th", "25th", "Median", "75th")
colors <- c(
  "5th" = "#4daf4a",
  "25th" = "#e41a1c",
  "75th" = "#377eb8",
  "Median" = "#000000"
)
color_scale_x <- list(
  scale_color_manual(
    "Pre",
    values = colors
  ),
    scale_fill_manual(
    "Pre",
    values = colors
  )
)
data.frame(
  pre = pre_values,
  label = pre_labels_x
  ) |> 
  kable(digits = 2)
```

Plot the interaction at different pretest values. As shown in @fig-lmm-pre-interaction, this interaction model shows a pattern similar to @fig-intercept-sds where those who report lower mood at the beginning of a session also improve more. 
```{r}
#| label: fig-lmm-pre-interaction
#| fig-cap: Interaction between pretest mood and change, shown at -2 SD, -1, SD, the mean, and +1 SD.
#| fig-height: 3
#| fig-width: 8
emm <- emmeans(
  fit3,
  ~ hours + pre,
  at = list(
    hours = seq(0, 3, length.out = 100),
    pre = pre_values
  ),
  lmer.df = "asymptotic"
)
p_pre_x <- as.data.frame(emm) |>
  mutate(
    pre = factor(
      pre,
      labels = pre_labels_x
      )
  ) |>
  ggplot(
    aes(
      hours,
      emmean,
      group = pre,
      color = pre
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      fill = pre,
      color = NULL
    ),
    alpha = 0.25,
    show.legend = FALSE
  ) +
  color_scale_x +
  scale_x_hours

p_pre_x
ggsave(
  "figures/fig-interaction-model.png", 
  p_pre_x, 
  width = 7, 
  height = 5
)
```

Let's also calculate the within-session change for the interaction model.
```{r}
#| label: fig-lmm-pre-interaction-change
#| fig-cap: Within-session change from the interaction model
#| fig-height: 2.5
#| fig-width: 8
source("R/get_change_by_pre.R")
d_x_change <- data.frame(
  pre = pre_values,
  label = pre_labels_x
) |>
  pmap(\(pre, label) get_change_by_pre(
    pre,
    label,
    fit3,
    hours = seq(0, 3, length.out = 100)
  )) |>
  bind_rows()

p_pre_x_change <- d_x_change |>
  mutate(
    label = factor(
      label,
      levels = pre_labels_x
    )
  ) |>
  ggplot(
    aes(
      hours,
      estimate,
      group = label,
      color = label,
      fill = label
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dotted"
  ) +
  labs(x = "Session duration (hours)") +
  color_scale_x +
  scale_x_hours +
  facet_wrap(~label, ncol = 4) 
p_pre_x_change
```

### Manuscript figures
Create figures for the manuscript.
```{r}
#| label: fig-ms-interaction
#| fig-height: 3.5
#| fig-width: 7
#| fig-cap: Figure included in manuscript

scale_x_hours2 <- list(
  scale_x_continuous(
  breaks = c(0, 15/60, 0.5, 45/60, 1, 2, 3), 
  labels = c(0, "", "30m", "", "1h", "2h", "3h")
  )
)
p_tmp <- (
  (
    p_pre_x + 
      facet_wrap(~pre, ncol = 4) +
      scale_x_hours2 +
      labs(y = "Mood") +
      theme(axis.title.x = element_blank())
  ) / 
    (
      p_pre_x_change +
        scale_x_hours2 +
        labs(y = "Change", x = "Duration") +
        theme()
    ) 
) +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "none") 
p_tmp
```

```{r}
#| label: ggsave-fig-ms-interaction
#| echo: false

ggsave(
  "figures/fig-pre-x-change.png", 
  p_tmp, 
  width = 7, 
  height = 3.5
  )
```

And this figure.

```{r}
#| label: fig-ms-no-interaction
#| fig-height: 3
#| fig-width: 6
#| fig-cap: Figure shown in manuscript

p_tmp <- (
  p_curve + labs(x = NULL, y = "Mood")
  ) / (
  p_change + labs(x = "Duration", y = "Change")
  ) + plot_annotation(tag_levels = "A") 
p_tmp
```

```{r}
#| label: ggsave-fig-ms-no-interaction
#| echo: false

ggsave(
  "figures/fig-change.png", 
  p_tmp, 
  width = 6, 
  height = 3
  )
```

And a table.
```{r}
#| label: tbl-pre-interaction-change
#| tbl-cap: Estimated within-session change from the interaction model
pre_sds <- c(1, 0, -1, -2)
tbl_x_change <- data.frame(
  pre = round(pre_values),
  label = pre_labels_x
) |>
  pmap(\(pre, label) get_change_by_pre(
    pre, 
    label, 
    fit3, 
    hours = c(0, 15/60, 0.5, 1, 2))
    ) |>
  bind_rows() |> 
  select(
    label, 
    pre, 
    hours,
    estimate, 
    CI_lwr = asymp.LCL, 
    CI_upr = asymp.UCL
    )
tbl_x_change |> 
  kable(digits = 2)
```

## Sensitivity checks
### Censoring

```{r}
#| label: n-pids-sample
#| echo: false
n_pids_sample <- case_when(
  is.na(N_SUBSET) ~ 1000,
  !is.na(N_SUBSET) ~ min(N_SUBSET, 1000)
)
```

The main discrete time model used a censored normal distribution. A similar model would take a long time to fit here, so we'll refit the continuous time model using a censored normal distribution on a random subset of `r n_pids_sample` participants.

```{r}
#| label: brms-cens
#| results: hide
set.seed(345345)
tmp <- filter(
  dat,
  pid %in% sample(unique(pid), n_pids_sample)
) |> 
  mutate(mood = mood / 100)

fit_b <- brm(
  mood | cens(cl) ~ ns4(hours) + (1 + ns4(hours) | pid) + (1 | ps),
  silent = 0,
  data = tmp,
  threads = 2,
  control = list(adapt_delta = .95),
  iter = 2000,
  prior = prior("normal(0,1)", class = "b"),
  file = "models/brms-ns4-cens"
)
```

We also re-fit the LMM to the reduced data.

```{r}
#| label: lmm-reduced
fit2 <- lmer(
  mood ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
  data = fit_b$data
)
```

If we plot the fitted spline from both models, we see in @fig-cens-compare that the shape is almost identical but the censored model is shifted upwards.

```{r}
#| label: fig-cens-compare
#| fig-cap: Estimated trends with and without censoring.
#| fig-height: 3
#| fig-width: 8
d_brms <-
  emmeans(
    fit_b,
    ~hours,
    at = list(hours = seq(0, 3, length.out = 100))
  ) |>
  as.data.frame() |>
  mutate(model = "cens") |>
  select(
    hours,
    emmean,
    model,
    lower = "lower.HPD",
    upper = "upper.HPD"
  )
d_lmm <-
  emmeans(
    fit2,
    ~hours,
    at = list(hours = seq(0, 3, length.out = 100)),
    lmer.df = "asymptotic"
  ) |>
  as.data.frame() |>
  select(
    hours,
    emmean,
    lower = "asymp.LCL",
    upper = "asymp.UCL"
  ) |>
  mutate(
    model = "lmm",
    emmean = emmean,
    upper = upper,
    lower = lower
  )
rbind(d_brms, d_lmm) |>
  ggplot(
    aes(
      hours,
      emmean,
      group = model,
      color = model
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = lower,
      ymax = upper,
      fill = model,
      color = NULL
    ),
    alpha = 0.25
  ) +
  labs(x = "Session duration (hours)")
```

Let's take a look at all model coefficients. @fig-model-cens-compare and @tbl-model-cens-compare show that ignoring censoring is perhaps not a huge problem.

```{r}
#| label: tidy-models
fit2_tidy <- tidy(fit2) |>
  mutate(
    model = "lmm",
    term = str_remove_all(term, "\\(|\\)")
  ) |>
  select(-statistic)
fit_b_tidy <- tidy(fit_b)
fit_b_tidy <- fit_b_tidy |>
  mutate(
    model = "cens",
    term = str_remove_all(term, "\\(|\\)")
  ) |>
  select(
    effect,
    group,
    term,
    estimate,
    std.error,
    model
  )
```

```{r}
#| label: fig-model-cens-compare
#| fig-cap: Model coefficients from an LMM with and without censoring
rbind(fit2_tidy, fit_b_tidy) |>
  ggplot(
    aes(
      y = term,
      estimate,
      color = model
    )
  ) +
  geom_point() +
  facet_wrap(
    ~effect,
    scale = "free",
    ncol = 1
  )
```

```{r}
#| label: tbl-model-cens-compare
#| tbl-cap: Model coefficients from an LMM with and without censoring
left_join(
  fit2_tidy,
  fit_b_tidy,
  by = join_by(term, group)
) |>
  mutate(
    diff = estimate.x - estimate.y,
    rel_diff = diff / estimate.y
  ) |>
  select(
    group,
    term,
    est_lmm = estimate.x,
    est_cens = estimate.y,
    diff,
    rel_diff
  ) |>
  kable(digits = 2)
```

We can also recreate @fig-intercept-sds using the Bayesian censored normal model and compare it to the LMM. As shown in @fig-bayes-int-sd, it's mostly the intercepts that are affected by the censoring.

```{r}
#| label: bayes-int-sd
source("R/get_trend_bayes.R")
res <- map(
  pre_sds,
  \(sd) get_trend_bayes(sd = sd)
) |>
  bind_rows() |>
  mutate(
    intercept = factor(
      sd,
      levels = pre_sds,
      labels = pre_labels
    )
  )

# lmm for comparison
res_lmm <- map(
  pre_sds,
  \(sd) get_trend(fit2, fit_b$data, sd = sd)
) |>
  bind_rows() |>
  mutate(
    intercept = factor(
      rep(
        pre_labels,
        each = 100
      ),
      levels = pre_labels
    )
  ) |>
  select(
    estimate = y,
    hours = x,
    intercept
  )

```

```{r}
#| label: fig-bayes-int-sd
#| fig-cap: Bayesian censored normal model vs LMM (dashed) fit to subset of data
#| fig-height: 3
#| fig-width: 8
ggplot(
  res,
  aes(
    hours, estimate,
    group = intercept,
    color = intercept
  )
) +
  geom_line() +
  geom_line(
    data = res_lmm,
    linetype = "dashed"
  ) +
  geom_ribbon(
    aes(
      x = hours,
      y = NULL,
      ymin = lwr,
      ymax = upr,
      color = NULL
    ),
    alpha = 0.15,
    show.legend = FALSE
  ) + 
  color_scale
```

Lastly, we'll also fit the interaction model using a censored normal distribution, with `r min(n_pids_sample, nlevels(dat_pre$pid))` participants.

```{r}
#| label: cens-pre-interaction
#| results: hide
set.seed(5454545)
tmp_pre <- filter(
    dat_pre,
    pid %in% sample(
        unique(pid),
        # avoid taking sample larger than population 
        min(
            n_pids_sample,
            nlevels(pid)
        )
    )
)

fit3_b <- brm(
  mood ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps),
  silent = 0,
  data = tmp_pre,
  threads = 2,
  #control = list(adapt_delta = .95),
  iter = 4000,
  file = "models/brms-ns4-interaction-cens"
)
```

Fit the LMM again to the smaller data.

```{r}
#| label: lmer-pre-interaction-small
 fit3_small <- lmer(
    mood ~ ns4(hours) * ns(pre, 5) + (1 + ns4(hours) | pid) + (1 | ps), 
    data = fit3_b$data
    )
```

We then compare the censored normal model with the LMM and recreate @fig-lmm-pre-interaction. As shown in @fig-cens-pre-interaction, there's almost no difference between the models.
```{r}
#| label: fig-cens-pre-interaction
#| fig-cap: Censored normal vs LMM with a continuous by continuous interaction between within-session change and pretest mood.
#| fig-height: 3
#| fig-width: 8
get_emm_pre <- function(fit) {
  emmeans(
    fit,
    ~ hours + pre,
    at = list(
      hours = seq(0, 3, length.out = 100),
      pre = c(
        pre_sum$mean - 2 * pre_sum$sd,
        pre_sum$mean - pre_sum$sd,
        pre_sum$mean,
        pre_sum$mean + pre_sum$sd
      )
    ),
    lmer.df = "asymp"
  )
}
pre_sum <- tmp_pre |>
  summarise(
    pre_mean = median(pre), .by = pid
  ) |>
  summarise(
    mean = mean(pre_mean), sd = sd(pre_mean)
  )
emm_cens <- get_emm_pre(fit3_b) |>
  as.data.frame() |>
  mutate(
    model = "cens"
  ) |>
  rename(
    lwr = lower.HPD,
    upr = upper.HPD
  )
emm_lmm <- get_emm_pre(fit3_small) |>
  as.data.frame() |>
  mutate(
    model = "LMM"
  ) |>
  rename(
    lwr = asymp.LCL,
    upr = asymp.UCL
  )

p_cens_pre_interaction <- bind_rows(emm_cens, emm_lmm) |>
  mutate(
    pre = factor(
      pre,
      levels = sort(unique(pre)),
      labels = c("-2 SD", "-1 SD", "Mean", "+1 SD")
    )
  ) |>
  {
    \(d)
    ggplot(
      data = d,
      aes(
        hours,
        emmean,
        group = interaction(pre, model),
        color = pre,
        linetype = model
      )
    ) +
      geom_line() +
      geom_ribbon(
        data = filter(d, model == "cens"),
        aes(
          ymin = lwr,
          ymax = upr,
          fill = pre,
          color = pre,
        ),
        color = NA,
        alpha = 0.25,
        show.legend = FALSE
      ) +
      geom_ribbon(
        data = filter(d, model == "LMM"),
        aes(
          ymin = lwr,
          ymax = upr,
          fill = pre,
          color = pre,
        ),
        fill = NA,
        alpha = 0.15,
        show.legend = FALSE
      ) +
      scale_color_brewer(palette = "Set1") +
      scale_fill_brewer(palette = "Set1")
  }()
p_cens_pre_interaction
```

### Functional form of `hours` and `pre`
Here we'll investigate the impact of the assumed functional form of `hours` and `pre`. As can be seen in @fig-lmm-pre-interaction-sens-pre, that impact of the assumed functional form of `pre` has very little impact on the results.

```{r}
#| label: fig-lmm-pre-interaction-sens-pre
#| fig-cap: Estimated trends for various functional forms of `pre`
tmp_models <- list()
tmp_models[["linear"]] <- fit_cached(
  "models/lmm-pre-interaction-sens-pre-linear.Rds",
  lmer(
    mood ~ ns4(hours) * pre + (1 + ns4(hours) | pid) + (1 | ps),
    data = dat_pre
  )
)
tmp_models[["ns2"]] <-
  fit_cached(
    "models/lmm-pre-interaction-sens-pre-ns2.Rds",
    lmer(
      mood ~ ns4(hours) * ns(pre, 2) + (1 + ns4(hours) | pid) + (1 | ps),
      data = dat_pre,
    )
  )
tmp_models[["ns3"]] <-
  fit_cached(
    "models/lmm-pre-interaction-sens-pre-ns3.Rds",
    lmer(
      mood ~ ns4(hours) * ns(pre, 3) + (1 + ns4(hours) | pid) + (1 | ps),
      data = dat_pre,
    )
  )
tmp_models[["ns4"]] <-
  fit_cached(
    "models/lmm-pre-interaction-sens-pre-ns4.Rds",
    lmer(
      mood ~ ns4(hours) * ns(pre, 4) + (1 + ns4(hours) | pid) + (1 | ps),
      data = dat_pre,
    )
  )
tmp_models[["ns5"]] <- fit3

get_trend_by_model <- function(fit, model_label) {
  emm <- emmeans(
    fit,
    ~ hours + pre,
    at = list(
      hours = seq(0, 3, length.out = 100),
      pre = pre_values
    ),
    lmer.df = "asymp"
  )
  as.data.frame(emm) |>
    mutate(
      pre = factor(
        pre,
        labels = pre_labels_x,
      ),
      model = model_label
    )
}
tmp_p <- map2(
  tmp_models,
  names(tmp_models),
  \(fit, lab) get_trend_by_model(fit, lab)
) |>
  bind_rows()
p_pre_x_sens <- tmp_p |>
  ggplot(
    aes(
      hours,
      emmean,
      group = interaction(model, pre),
      color = model,
      linetype = pre
    )
  ) +
  geom_line() +
  geom_ribbon(
    aes(
      ymin = asymp.LCL,
      ymax = asymp.UCL,
      fill = model,
      color = NULL
    ),
    alpha = 0.10,
    show.legend = FALSE
  ) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  scale_x_hours
p_pre_x_sens
```

Next, we'll look at `hours`. In @fig-lmm-pre-interaction-sens-hours we see that the results aren't fundamentally different using splines with less degrees of freedom, or even a linear term.

```{r}
#| label: fig-lmm-pre-interaction-sens-hours
#| fig-cap: Estimated trends for various functional forms of `hours`  
tmp_models2 <- list()
tmp_models2[["linear"]] <- fit_cached(
  "models/lmm-pre-interaction-sens-hours-linear.Rds",
  lmer(
    mood ~ hours * ns(pre, 5) + (1 + hours | pid) + (1 | ps),
    data = dat_pre
  )
)
tmp_models2[["ns2"]] <-
  fit_cached(
    "models/lmm-pre-interaction-sens-hours-ns2.Rds",
    lmer(
      mood ~ ns(hours, 2) * ns(pre, 5) + (1 + ns(hours, 2) | pid) + (1 | ps),
      data = dat_pre,
    )
  )
tmp_models2[["ns3"]] <- fit_cached(
  "models/lmm-pre-interaction-sens-hours-n3.Rds",
  lmer(
    mood ~ ns(hours, 3) * ns(pre, 5) + (1 + ns(hours, 3) | pid) + (1 | ps),
    data = dat_pre,
  )
)
tmp_models2[["poly2"]] <- fit_cached(
  "models/lmm-pre-interaction-sens-hours-poly2.Rds",
  lmer(
    mood ~ poly(hours, 2) * ns(pre, 5) + (1 + poly(hours, 2) | pid) + (1 | ps),
    data = dat_pre,
  )
)
tmp_models2[["ns4"]] <- fit3
tmp_p <- map2(
  tmp_models2,
  names(tmp_models2),
  \(fit, lab) get_trend_by_model(fit, lab)
) |>
  bind_rows()
p_pre_x_sens %+% tmp_p 
```